---
title: "Data Analytics code"
output: html_document
date: "2023-09-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(gbm)
library(xgboost)
library(ipred)
library(pdp)
library(glmnet)
library(broom)
library(readxl)
library(randomForest)
library(caret)
library(gam)
library(mgcv)
library(earth)
library(rpart)
library(tidymodels)
library(rsample)
library(ggplot2)
library(coefplot)
```

Loading & Cleaning the Data 
```{r}
cleaned_data <- read_excel('/Users/vigneshravikumar/Desktop/DA Project/cleaned_data1.xlsx')
cleaned_data<- unique(cleaned_data)
cleaned_data <- na.omit(cleaned_data)
```

**Spliting the Data**
```{r}
split_ratio <- 0.8
train_indices <- createDataPartition(cleaned_data$Weekly_Sales, p = split_ratio, list = FALSE)
trainingdata <- cleaned_data[train_indices, ]
testdata <- cleaned_data[-train_indices, ]

```
names(cleaned_data)
```{r}
library(corrplot)
corr_data= subset(cleaned_data,select = c(Weekly_Sales,IsHoliday,Temperature,Fuel_Price,MarkDown1,MarkDown2,MarkDown3,MarkDown4,MarkDown5,CPI,Unemployment,Size))

correlation_matrix <- cor(corr_data)
corrplot(correlation_matrix, method = "color")
```


**Multiple Linear Regression model**
```{r}
mlr_model <- lm(Weekly_Sales ~ Temperature+ Fuel_Price + MarkDown1 + MarkDown2 + MarkDown3 + MarkDown4 + MarkDown5 + CPI + Unemployment + IsHoliday + Size + Type, 
                data = trainingdata) # fit

mlr_ypred <- predict(mlr_model, testdata) # predict

# Training R^2
print(paste('Training R^2: ', summary(mlr_model)$r.squared, sep = ''))

# Test R^2
print(paste('Test R^2: ', cor(testdata$Weekly_Sales, mlr_ypred)^2, sep = ''))

# Training RMSE
print(paste('Training RMSE: ','$', 
            sqrt(mean((trainingdata$Weekly_Sales - mlr_model$fitted.values)^2)), 
             sep = ''))

# Test RMSE
print(paste('Test RMSE: ','$', 
            sqrt(mean((testdata$Weekly_Sales - mlr_ypred)^2)), 
             sep = ''))
```

```{r}
model_summary<-summary(mlr_model)
model_summary
```
```{r}
p_values <- model_summary$coefficients[, "Pr(>|t|)"]
variable_names <- rownames(model_summary$coefficients)

p_value_table <- data.frame(Variable = variable_names, P_Value = p_values)

sorted_table <- p_value_table[order(p_value_table$P_Value), ]

print(sorted_table)
```
```{r}
p_value_plot_data <- data.frame(Variable = variable_names, P_Value = p_values)
ggplot(p_value_plot_data, aes(x = Variable, y = P_Value)) +
  geom_point(size = 3, color = "lightblue") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "p-values of predictor variables wrt Weekly_Sales",
       x = "predictor variables",
       y = "p-values")
```


</br>
We find the predictor variables using the MLR model summary. The predictor variables that we have chosen are MarkDown1, MarkDown3, CPI,Unemployment,IsHoliday,Size,Type. The Response Variable is Weekly Sales
</br>


**Model 1 GLM Model**
```{r}
glm_model <- glm(Weekly_Sales~ Fuel_Price + MarkDown3 +MarkDown4 + CPI +Unemployment + Size+ Type, data = trainingdata) 
glm_ypred <- predict(glm_model, testdata) 



# Test R^2
print(paste('Test R^2: ', cor(testdata$Weekly_Sales, glm_ypred)^2, sep = ''))

# Training RMSE
print(paste('Training RMSE: ','$', 
            sqrt(mean((trainingdata$Weekly_Sales - glm_model$fitted.values)^2)), 
             sep = ''))
print(paste('Test RMSE: ','$', sqrt(mean((testdata$Weekly_Sales - glm_ypred)^2))))
```
```{r}
ggplot(cleaned_data, aes(y = Weekly_Sales)) +
  geom_boxplot() +
  labs(title = "Boxplot of Weekly Sales", y = "Weekly Sales")

```
```{r}
coef_data <- tidy(glm_model)

# Create a coefficient plot
coefplot(
  coef_data,
  title = "Coefficient Plot for GLM",
  ylab = "Predictor",
  interval = "confidence",
  errorbar = "x",  # Specify the error bars in the horizontal direction
  xlim = c(-2, 2)  # Adjust the x-axis limits as needed
)
```


```{r}
# Assuming you have a "Date" column in your testdata dataframe
# If your "Date" column is not in Date format, convert it to Date format
testdata$Date <- as.Date(testdata$Date)

# Create a new dataframe with Date, actual values, and predicted values
comparison_df <- data.frame(
  Date = testdata$Date,
  Actual = testdata$Weekly_Sales,
  Predicted = glm_ypred
)

# Load necessary libraries
library(ggplot2)

# Create a line graph
plot <- ggplot(data = comparison_df, aes(x = Date)) +
  geom_line(aes(y = Actual, color = "Actual"), size = 1) +
  geom_line(aes(y = Predicted, color = "Predicted"), size = 1, linetype = "dashed") +
  labs(x = "Date", y = "Weekly Sales", color = "Legend") +
  scale_color_manual(values = c("Actual" = "blue", "Predicted" = "red"), labels = c("Actual", "Predicted")) +
  ggtitle("Comparison of Actual and Predicted Weekly Sales GLM model")

# Set the width and height (in inches) to increase the total size
width_inch <- 30  # Set your desired width
height_inch <- 4   # Set your desired height

# Save the plot as an image file with the specified width and height
ggsave("GLMcomparison_plot.png", plot, width = width_inch, height = height_inch)
```

**Model 2 GAM Model**
```{r}
gam_model <- gam( Weekly_Sales ~  Fuel_Price + MarkDown2 +MarkDown3 + CPI +Unemployment +IsHoliday+ Size+ Type,
              data = trainingdata)

gam_ypred <- predict(gam_model, newdata = testdata)
print(paste('Test RMSE: ','$',sqrt(mean((testdata$Weekly_Sales - gam_ypred)^2, na.rm = T))))
```
```{r}
# Assuming you have a "Date" column in your testdata dataframe
# If your "Date" column is not in Date format, convert it to Date format
testdata$Date <- as.Date(testdata$Date)

# Create a new dataframe with Date, actual values, and predicted values
comparisongam_df <- data.frame(
  Date = testdata$Date,
  Actual = testdata$Weekly_Sales,
  Predicted = gam_ypred
)

# Load necessary libraries
library(ggplot2)

# Create a line graph
plot <- ggplot(data = comparisongam_df, aes(x = Date)) +
  geom_line(aes(y = Actual, color = "Actual"), size = 1) +
  geom_line(aes(y = Predicted, color = "Predicted"), size = 1, linetype = "dashed") +
  labs(x = "Date", y = "Weekly Sales", color = "Legend") +
  scale_color_manual(values = c("Actual" = "blue", "Predicted" = "red"), labels = c("Actual", "Predicted")) +
  ggtitle("Comparison of Actual and Predicted Weekly Sales GAM model")

# Set the width and height (in inches) to increase the total size
width_inch <- 30  # Set your desired width
height_inch <- 4   # Set your desired height

# Save the plot as an image file with the specified width and height
ggsave("Gamcomparison_plot.png", plot, width = width_inch, height = height_inch)
```

**Model 3 MARS Model**
```{r}
mars <- earth(Weekly_Sales ~  Fuel_Price + MarkDown2 +MarkDown3 + CPI +Unemployment +IsHoliday+ Size+ Type, 
                data = trainingdata)
mars_ypred <- predict(mars, testdata)
# Assuming you have a dataframe named 'df' and you want to change the column name 'OldName' to 'NewName'
colnames(mars_ypred)[colnames(mars_ypred) == 'Weekly_Sales'] <- 'Predicted'
print(paste('Test RMSE: ','$',sqrt(mean((testdata$Weekly_Sales - mars_ypred)^2, na.rm = T))))
```
```{r}
# Assuming you have a "Date" column in your testdata dataframe
# If your "Date" column is not in Date format, convert it to Date format
testdata$Date <- as.Date(testdata$Date)

# Create a new dataframe with Date, actual values, and predicted values
comparisonmars_df <- data.frame(
  Date = testdata$Date,
  Actual = testdata$Weekly_Sales,
  Predicted = mars_ypred
)

# Load necessary libraries
library(ggplot2)

# Create a line graph
plot <- ggplot(data = comparisonmars_df, aes(x = Date)) +
  geom_line(aes(y = Actual, color = "Actual"), size = 1) +
  geom_line(aes(y = Predicted, color = "Predicted"), size = 1, linetype = "dashed") +
  labs(x = "Date", y = "Weekly Sales", color = "Legend") +
  scale_color_manual(values = c("Actual" = "blue", "Predicted" = "red"), labels = c("Actual", "Predicted")) +
  ggtitle("Comparison of Actual and Predicted Weekly Sales MARS model")

# Set the width and height (in inches) to increase the total size
width_inch <- 30  # Set your desired width
height_inch <- 4   # Set your desired height

# Save the plot as an image file with the specified width and height
ggsave("MARScomparison_plot.png", plot, width = width_inch, height = height_inch)
```

**Model 4 CART Model**
```{r}
cart_model <- rpart(Weekly_Sales ~  Fuel_Price + MarkDown2 +MarkDown3 + CPI +Unemployment +IsHoliday+ Size+ Type, data = trainingdata)
                
cart_ypred2 <- predict(cart_model, testdata)

rmse <- sqrt(mean((testdata$Weekly_Sales - cart_ypred2)^2))
print(paste("Test RMSE","$",rmse))
```
```{r}
# Assuming you have a "Date" column in your testdata dataframe
# If your "Date" column is not in Date format, convert it to Date format
testdata$Date <- as.Date(testdata$Date)

# Create a new dataframe with Date, actual values, and predicted values
comparisoncart_df <- data.frame(
  Date = testdata$Date,
  Actual = testdata$Weekly_Sales,
  Predicted = cart_ypred2
)

# Load necessary libraries
library(ggplot2)

# Create a line graph
plot <- ggplot(data = comparisoncart_df, aes(x = Date)) +
  geom_line(aes(y = Actual, color = "Actual"), size = 1) +
  geom_line(aes(y = Predicted, color = "Predicted"), size = 1, linetype = "dashed") +
  labs(x = "Date", y = "Weekly Sales", color = "Legend") +
  scale_color_manual(values = c("Actual" = "blue", "Predicted" = "red"), labels = c("Actual", "Predicted")) +
  ggtitle("Comparison of Actual and Predicted Weekly Sales MARS model")

# Set the width and height (in inches) to increase the total size
width_inch <- 30  # Set your desired width
height_inch <- 4   # Set your desired height

# Save the plot as an image file with the specified width and height
ggsave("Cartcomparison_plot.png", plot, width = width_inch, height = height_inch)
```

**Model 5 Bagging Model**
```{r}
bagging_mod <- bagging(Weekly_Sales ~  Fuel_Price + MarkDown2 +MarkDown3 + CPI +Unemployment +IsHoliday+ Size+ Type, data = trainingdata, nbagg = 100, coob = TRUE)
ypred <- predict(bagging_mod, newdata = testdata)


print(paste('Test RMSE: ','$', round(sqrt(mean((testdata$Weekly_Sales - ypred)^2)),2)))
```

**Model 6 Random Forests**
```{r}
rf_mod <- randomForest(Weekly_Sales ~  Fuel_Price + MarkDown2 +MarkDown3 + CPI +Unemployment +IsHoliday+ Size+ Type, 
                       data = trainingdata, ntree = 100, 
                       mtry = 3, nodesize = 10,
                       importance = TRUE)
ypred_rf <- predict(rf_mod, newdata = testdata)

print(paste('Test RMSE: ','$', round(sqrt(
  mean((testdata$Weekly_Sales - ypred_rf)^2)), 2)))
```



**Model 7 Boosting Model**
```{r}
gbm_mod <- gbm(Weekly_Sales ~  Fuel_Price + MarkDown2 +MarkDown3 + CPI +Unemployment +IsHoliday+ Size+ Type, data = trainingdata, n.trees = 500,shrinkage = 0.01, interaction.depth = 3)
ypred_gbm <- predict(gbm_mod, testdata, n.trees = 500) 


print(paste('Test RMSE: ','$',round(sqrt(mean((testdata$Weekly_Sales - ypred_gbm)^2)),2)))
```

## K-Fold Method
**Model 1 5-fold GLM Model**
```{r}
kfold_fold <- 5
folds <- cut(seq(1,nrow(cleaned_data)),breaks=kfold_fold,labels=FALSE)
sum <- 0
for(i in 1:kfold_fold){
    testIndexes <- which(folds==i,arr.ind=TRUE)
    test_data <- cleaned_data[testIndexes, ]
    training_data <- cleaned_data[-testIndexes, ]
    
    glm_model <- glm( Weekly_Sales ~  Fuel_Price + MarkDown2 +MarkDown3 + CPI +Unemployment +IsHoliday+ Size+ Type, data = training_data, family = gaussian(link = 'identity')) 

    glm_ypred <- predict(glm_model, test_data) 
    rmse_value <- sqrt(mean((test_data$Weekly_Sales - glm_ypred)^2))
sum <- sum + rmse_value
}
avg_rmse <- sum/kfold_fold
print(paste('GLM Avg Test RMSE: ',  '$', avg_rmse, sep = ''))
```

**Model 2 5-fold GAM Model**
```{r}
kfold_foldgam <- 5
foldsgam <- cut(seq(1,nrow(cleaned_data)),breaks=kfold_foldgam,labels=FALSE)
sum <- 0
for(i in 1:kfold_foldgam){
    testIndexes <- which(foldsgam==i,arr.ind=TRUE)
    test_data <- cleaned_data[testIndexes, ]
    training_data <- cleaned_data[-testIndexes, ]
    gam_model <- gam(Weekly_Sales ~ s(Fuel_Price) + s(MarkDown2) +s(MarkDown3)+ s(CPI) + s(Unemployment) + IsHoliday + s(Size) + Type, data = training_data) 
gam_ypred <- predict(gam_model, test_data) 
rmsegam_value <- sqrt(mean((test_data$Weekly_Sales - gam_ypred)^2))
sum <- sum + rmsegam_value
}
avg_rmse <- sum/kfold_foldgam
print(paste('GAM Avg Test RMSE: ','$',avg_rmse, sep = ''))
```

**Model 3 5-fold MARS Model**
```{r}
kfold_fold <- 5
folds <- cut(seq(1,nrow(cleaned_data)),breaks=kfold_fold,labels=FALSE)
sum <- 0
for(i in 1:kfold_fold)
  {
    testIndexes <- which(folds==i,arr.ind=TRUE)
    testdata <- cleaned_data[testIndexes, ]
    trainingdata <- cleaned_data[-testIndexes, ]
    mars_model <- earth (Weekly_Sales ~  Fuel_Price + MarkDown2 +MarkDown3 + CPI +Unemployment +IsHoliday+ Size+ Type,
                         data = trainingdata) 
mars_ypred <- predict(glm_model, testdata) 
rmse_value <- sqrt(mean((testdata$Weekly_Sales - mars_ypred)^2))
sum <- sum + rmse_value
}
avg_rmse <- sum/kfold_fold
print(paste('MARS Avg Test RMSE: ', '$', avg_rmse, sep = ''))
```

**Model 4 5-fold CART Model**
```{r}
set.seed(123)
k <- 5
n <- nrow(cleaned_data)
fold_indices <- sample(1:k, n, replace = TRUE)
test_rmse_values <- numeric(k)
for (i in 1:k) 
  {
  train_data <- cleaned_data[fold_indices != i, ]
  test_data <- cleaned_data[fold_indices == i, ]

  cart_model <- rpart(Weekly_Sales ~ Fuel_Price + MarkDown2 +MarkDown3 + CPI +Unemployment +IsHoliday+ Size+ Type, data = train_data)

  test_predictions <- predict(cart_model, newdata = test_data)
  test_rmse_values[i] <- sqrt(mean((test_data$Weekly_Sales - test_predictions)^2))
}
average_test_rmse <- mean(test_rmse_values)
print(paste("Average Test RMSE:",'$', average_test_rmse))
```


**Model 5 5-fold Bagging Cross Validations**
```{r}
set.seed(123)
k <- 5
n <- nrow(cleaned_data)
fold_indices <- sample(1:k, n, replace = TRUE)
test_rmse_values <- numeric(k)
for (i in 1:k) 
  {
  train_data <- cleaned_data[fold_indices != i, ]
  test_data <- cleaned_data[fold_indices == i, ]

  bagging_mod <- bagging(Weekly_Sales ~  Fuel_Price + MarkDown2 +MarkDown3 + CPI +Unemployment +IsHoliday+ Size+ Type,
                         data = train_data, nbagg = 100, coob = TRUE)

  test_predictions <- predict(bagging_mod, newdata = test_data)
  test_rmse_values[i] <- sqrt(mean((test_data$Weekly_Sales - test_predictions)^2))
}
average_test_rmse <- mean(test_rmse_values)
print(paste("Average Test RMSE: ",'$', average_test_rmse))

```

**Model 6 5-Fold Boosting Method**
```{r}
gbm_mod2 <- gbm(Weekly_Sales ~ Fuel_Price + MarkDown2 +MarkDown3 + CPI +Unemployment +IsHoliday+ Size+ Type, data = trainingdata, n.trees = 2000, shrinkage = 0.2, interaction.depth = 3, cv.folds = 5)
ypred_gbm <- predict(gbm_mod2, testdata, n.trees = 500) # need to specify the number of trees -- same as training model

print(paste('Test RMSE: ','$', round(sqrt(mean((testdata$Weekly_Sales - ypred_gbm)^2)),2)))
```


**Model 7 5-fold Hypertuned MARS Model**
```{r}
cv <- vfold_cv(cleaned_data, v = 5)

model_recipe <- recipe(Weekly_Sales ~  Fuel_Price + MarkDown2 +MarkDown3 + CPI +Unemployment +IsHoliday+ Size+ Type, data= cleaned_data)

mars_mod5 <- mars(num_terms = tune(), prod_degree = tune(), 
                  prune_method = tune()) %>% set_engine('earth') %>%
  set_mode('regression')

mars_wf <- workflow() %>% add_recipe(model_recipe) %>% add_model(mars_mod5)

mars_grid <- expand.grid(
  num_terms = c(8,10,12,14,16),
  prod_degree =1:4 ,
  prune_method = c('backward', 'forward', 'none')
)
mars_tuning_results <- mars_wf %>% tune_grid(resamples = cv, grid = mars_grid)
mars_best_parameters <- mars_tuning_results %>% select_best(metric = 'rmse')
mars_wf2 <- mars_wf %>% finalize_workflow(mars_best_parameters)
mars_fit <- mars_wf2 %>% fit_resamples(cv, 
                                       control = control_resamples(
                                         save_pred = T))
performance <- mars_fit %>% collect_metrics()
print(paste('Tuned MARS Model Test RMSE: ','$', performance[1,3], sep = ''))

```

**Model 8 5-fold RandomForest**
```{r}
kfold_fold <- 5
folds <- cut(seq(1,nrow(cleaned_data)),breaks=kfold_fold,labels=FALSE)
sum <- 0
for(i in 1:kfold_fold){
    testIndexes <- which(folds==i,arr.ind=TRUE)
    test_data <- cleaned_data[testIndexes, ]
    training_data <- cleaned_data[-testIndexes, ]
rf_mod <- randomForest(Weekly_Sales ~  Fuel_Price + MarkDown2 +MarkDown3 + CPI +Unemployment +IsHoliday+ Size+ Type, 
                       data = training_data, ntree = 100, 
                       mtry = 3, nodesize = 10,
                       importance = TRUE)
ypred_rf <- predict(rf_mod, newdata = test_data)
    rmse_value <- sqrt(mean((test_data$Weekly_Sales - ypred_rf)^2))
sum <- sum + rmse_value
}
avg_rmse <- sum/kfold_fold
print(paste('RF Avg Test RMSE: $ ', avg_rmse, sep = ''))
```

```{r}
RMSEval <- read_excel("C:/Users/chara/Documents/Documents/EME 597 Data Analytics/Project/Final RMSE values.xlsx")
```
```{r}

# Create a bar plot
ggplot(data = RMSEval, aes(x = Model, y = RMSE)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Supervised Learning Models", y = "Test RMSE") +
  ggtitle("Test RMSE for Supervised Learning Models")

```

